---
title: "geovis_assignment2"
output: html_document
date: "2025-05-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load necessary libraries
```{r load}

library(sf)        # For handling spatial (vector) data

library(ggplot2)   # For creating elegant plots 
# install ggplot2 if using it the first time
#install.packages ("ggplot2")
```

### Step 1: Load pollution data from github (online source)
```{r}

pollution_data <- read.csv("https://raw.githubusercontent.com/mengluchu/uncertainty/master/data_vis_exp/DENL17_uc.csv")

```

### Step 2: Convert the table to a spatial 'sf' object using longitude and latitude
```{r}

sf_pollution <- st_as_sf(pollution_data, coords = c("Longitude", "Latitude"), crs = 4326)  # EPSG:4326 = WGS84

```
### Step 3: Check the data structure

```{r}
head(sf_pollution)
```

### Step 4: Create a Map using ggplot2

```{r }

#We will now visualize:  NO2 values with color
#Population density with size

ggplot(sf_pollution) +
  geom_point(aes(x = st_coordinates(geometry)[, 1],
                 y = st_coordinates(geometry)[, 2],
                 color = wkd_day_value,
                 size = population_1000),
             alpha = 0.6) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Weekday NO2 (Daytime) vs Population",
       x = "Longitude", y = "Latitude",
       color = "NO2_daytime_weekday_2017",
       size = "Population density")


```
### Step 5
Adjust Symbol Size (e.g. for Population)
 You can represent a numeric variable (like population) using symbol size:
 aes(size = population_1000)   # larger values means bigger points
 You can further control the size range by adding
 scale_size(range = c(1, 10))  # sets minimum and maximum point sizes

```{r}


ggplot(sf_pollution) +
  geom_point(aes(x = st_coordinates(geometry)[, 1],
                 y = st_coordinates(geometry)[, 2],
                 color = wkd_day_value,
                 size = population_1000),
             alpha = 0.6) +
  scale_size(range = c(1, 10)) +
  scale_color_viridis_c() +
  theme_minimal() +
  labs(title = "Weekday NO2 (Daytime) vs Population",
       x = "Longitude", y = "Latitude",
       color = "NO2_daytime_weekday_2017",
       size = "Population density")


```

### Step 6
 you can also control the colour, for example, specify the "option" in
scale_color_viridis_c(option = "plasma")  # other options: "magma", "inferno", "cividis"

```{r}

ggplot(sf_pollution) +
  geom_point(aes(x = st_coordinates(geometry)[, 1],
                 y = st_coordinates(geometry)[, 2],
                 color = wkd_day_value,
                 size = population_1000),
             alpha = 0.6) +
  scale_color_viridis_c(option = "plasma") +
  theme_minimal() +
  labs(title = "Weekday NO2 (Daytime) vs Population",
       x = "Longitude", y = "Latitude",
       color = "NO2_daytime_weekday_2017",
       size = "Population density")

 
```

### Step 7
 Adjust Point Shape and Transparency, for example, add in the geom_point the "shape " and adjust "alpha":
 geom_point(shape = 21, alpha = 0.6)  # semi-transparent circles

```{r}

 ggplot(sf_pollution ) +
  geom_point(aes(x = st_coordinates(geometry)[, 1],
                 y = st_coordinates(geometry)[, 2],
                 color = wkd_day_value,
                 size = population_1000), 
             shape = 21,
             alpha = 0.6) +
  scale_color_viridis_c(option = "plasma") +
  theme_minimal() +
  labs(title = "Weekday NO2 (Daytime) vs Population",
       x = "Longitude", y = "Latitude",
       color = "NO2_daytime_weekday_2017",
       size = "Population density")
 

```
### Optional: Step 8 add a scale bar
```{r}

# install.packages("ggspatial")
 library(ggspatial)
 sf_projected <- st_transform(sf_pollution, crs = 32632)  #project it so that the scale is in km or m. UTM Zone 32N for western Germany and 33N (epsg: 32633) for eastern Germany
 
 ggplot(sf_projected) +
   geom_point(aes(x = st_coordinates(geometry)[, 1],
                  y = st_coordinates(geometry)[, 2],
                  color = wkd_day_value,
                  size = population_1000), 
              shape = 21,
              alpha = 0.6) +
   scale_color_viridis_c(option = "plasma") +
   theme_minimal() +
   labs(title = "Weekday NO2 (Daytime) vs Population",
        x = "Longitude", y = "Latitude",
        color = "NO2_daytime_weekday_2017",
        size = "Population density")+
   annotation_scale(location = "bl", width_hint = 0.5) +  # Bottom left
   annotation_north_arrow(location = "bl", which_north = "true")

 #Export your result, use
 #ggsave("PATH/filename.png"), PATH = where you want to store your file. filename = your filename, ".png".  
 #see https://ggplot2.tidyverse.org/reference/ggsave.html for further configurations.

```

 