---
title: "geovis_L1"
output: html_document
date: "2025-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("sf","spData")

```
We will use the simple feature package.

Simple features or simple feature access refers to a formal standard (ISO 19125-1:2004) that describes how objects in the real world can be represented in computers, with emphasis on the spatial geometry of these objects. 
An hierarchical data model that represents a wide range of geometry types. 

```{r}
library(sf)
```


An sf object contains the following objects of class sf, sfc and sfg:
* sf (simple feature): each row of the data.frame is a single simple feature consisting of attributes and geometry.
* sfc (simple feature geometry list-column): the geometry column of the data.frame is a list-column of class sfc with the geometry of each simple feature.
* sfg (simple feature geometry): each of the rows of the sfc list-column corresponds to the simple feature geometry (sfg) of a single simple feature.

If you are curious about the simple feasure, you can have a read at the R document
https://r-spatial.github.io/sf/articles/sf1.html
wikipidia, or lecture slides I made for another course:
https://docs.google.com/presentation/d/1cJUOpMnA84nnOzgh8GvoKJaMttOl3xyqQVubKydYKfg/edit?usp=sharing

#### Dataset world
Use the **world** dataset provided by the **spData**,  see [nowosad.github.io/spData] for a list of datasets loaded by the package.
```{r}
#install.packages("spData")
library(spData)
```
 **world** is a spatial object containing spatial and attribute columns.
 so_a2 character vector of ISO 2 character country codes

* name_long character vector of country names

* continent character vector of continent names

* region_un character vector of region names

* subregion character vector of subregion names

* type character vector of type names

* area_km2 integer vector of area values

* pop integer vector of population in 2014

* lifeExp integer vector of life expectancy at birth in 2014

* gdpPercap integer vector of per-capita GDP in 2014

* geom sfc_MULTIPOLYGON

```{r summary}
summary(world) # tells you the statistics of data, the variables
```
 

```{r head}
head(world) # shows the first few rows of code to explore.
```


```{r plot}
plot(world["lifeExp"])
```
Plot the life expectancy by (a different) interval
```{r}
breaks <- c(50, 60, 70, 80 )
plot(world["lifeExp"], breaks = breaks)

```


```{r}
hist(world["lifeExp"]$lifeExp)
```

Find the number of missing values for each feature
```{r}
colSums(is.na(world))

```
Compare life expectancy between continent
```{r}
boxplot(lifeExp ~ continent, data = world,
        main = "Life Expectancy by Continent",
        xlab = "Continent", ylab = "Life Expectancy")

```

Advanced: a professional tool for mapping and interactive exploration: tmap
```{r}
#install.packages("tmap")
library(tmap)
```
Explore with the interactive and static map. 
```{r}
tmap_mode("view") #interactive
#tmap_mode("plot") # for static map 
```

You can change the transparency (alpha).
```{r}

tm_shape(world) +
  tm_polygons("lifeExp", 
              style = "quantile", 
              palette = "Blues", 
              title = "Life Expectancy",
              alpha = 0.5) +
  tm_layout(title = "World Life Expectancy Map")
```
 
 